---
title: "Project 1 Checkpoint 2"
author: "Scarlet Kwan"
format:
  html: 
    theme: darkly
    code-fold: show
    code-tools: true
    embed-resources: true
editor: source
execute: 
  eval: true
  echo: true
  warning: false
---

```{r}

democracy_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-11-05/democracy_data.csv')
library(tidyverse)
library(gt)

```


## Data Summaries Unmodified

**Research Questions Based on Democracy Data:**

- Are presidential systems more likely to be democratic?
 
```{r}
democracy_presid_voting <- function(df) {
  df |>
    group_by(is_presidential) |>
    summarize(
      prop_democracy = mean(is_democracy, na.rm = TRUE),
      count = n()) |>
    arrange(desc(prop_democracy))
}

democracy_presid_voting(democracy_data)
```

-   Does the inclusion of proportional voting increase the possibility of a free and fair election?

```{r}
prop_vs_fair_map <- democracy_data |>
  split(~has_proportional_voting) |>
  map_dbl(~ mean(.x$has_free_and_fair_election, na.rm = TRUE))

```


```{r}
library(tibble)

prop_vs_fair <- prop_vs_fair_map |>
  enframe(name = "proportional voting", value = "prop of free & fair")|> gt::gt()

prop_vs_fair
```



## Data Summaries Modified

- How many female president or monarch leaders have led a country over time?


```{r}
#custom function w/ new column

summarize_female_leadership <- function(data) {
   data |>
    mutate(
      female_leader = is_female_monarch == TRUE | is_female_president == TRUE
    ) |>
    group_by(year) |>
    summarize(
      number_of_countries_per_year = n(),
      female_leaders = sum(female_leader, na.rm = TRUE),
      proportion_female = female_leaders / number_of_countries_per_year
    )
}
```

```{r}
#using the function + gt table
summarize_female_leadership(democracy_data) |>
  gt() |>
  tab_header(
    title = "Female Leadership Over Time",
    subtitle = "Proportion of Female Leaders by Year"
  ) |>
  cols_label(
    year = "Year",
    number_of_countries_per_year = "Total Countries In A Year",
    female_leaders = "Female Leaders",
    proportion_female = "Proportion Female"
  ) |>
  fmt_percent(
    columns = proportion_female,
    decimals = 2
  )
```



- How many colonies are present over time?

```{r}
democracy_data <- democracy_data |>
  mutate(
    colony_status = if_else(is_colony == TRUE, "Colony", "Not Colony")
  )
```
 
```{r}
# this function filters the number of colunies in a specific inputted year
num_colonies <- function(y) {
  democracy_data |>
    filter(year == y, colony_status == "Colony") |>
    nrow()
}
```


```{r}
#using the above functions, we are able to map the variable number_of_colonies for each year
colony_summary <- democracy_data |>
  distinct(year) |>
  arrange(year) |>
  mutate(
    number_of_colonies = map_int(year, num_colonies),
    proportion_of_colonies = number_of_colonies / 208
  )

#gt table here

colony_summary |>
  gt() |>
  tab_header(
    title = "Chnage In Number of Colonies by Year",
    subtitle = "208 Countries In Our Dataset") |>
  cols_label(
    year = "Year",
    number_of_colonies = "Number of Colonies",
    proportion_of_colonies = "Proportion of Countries That Are Colonies",
    
  ) |>
  fmt_percent(
    columns = proportion_of_colonies,
    decimals = 2
  )
```