---
title: "Project 1 Checkpoint 2"
author: "Scarlet Kwan"
format:
  html: 
    theme: darkly
    code-fold: show
    code-tools: true
    embed-resources: true
editor: source
execute: 
  eval: true
  echo: true
  warning: false
---

## Data Summaries Unmodified

<<<<<<< HEAD
```{r}
#| output: false
democracy_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-11-05/democracy_data.csv')
```

**Research Questions Based on Democracy Data:**

- Are presidential systems more likely to be democratic?
 
```{r}
democracy_presid_voting <- function(df) {
  df |>
    group_by(is_presidential) |>
    summarize(
      prop_democracy = mean(is_democracy, na.rm = TRUE),
      count = n()) |>
    arrange(desc(prop_democracy))
}

democracy_presid_voting(democracy_data)
```

-   Does the inclusion of proportional voting increase the possibility of a free and fair election?

```{r}
library(dplyr)
library(purrr)

prop_vs_fair_map <- democracy_data |>
  split(~has_proportional_voting) |>
  map_dbl(~ mean(.x$has_free_and_fair_election, na.rm = TRUE))

```


```{r}
library(tibble)

prop_vs_fair <- prop_vs_fair_map |>
  enframe(name = "proportional voting", value = "prop of free & fair")|> gt::gt()

prop_vs_fair
```

## Data Summaries Modified

Description
